Computational design of RNA-based oscillatory circuits

### 2015-08-30 MScWriteup.tex
rewrites to the intro

\section{Introduction}
\label{sec: Intro}
The process of gene expression can be briefly summarised as: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}). This RNA molecule (known as messenger RNA, or mRNA) makes its way to a piece of cellular machinery called the Ribosome, which reads it, and makes a protein - which protein is made depends on the DNA sequence originally read (\textit{translation}).

 The path from genetic transcription to protein expression is naturally regulated in many ways \cite{MolecularBiology}. This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues. The natural cell machinery which performs this regulation offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.

As well as acting as the intermediate between DNA and protein, RNA molecules play direct and important roles in regulating cell behaviour \cite{Isaacs2006}. For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA's offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription. In comparison to proteins, it is relatively straightforward to predict the structure and function of an RNA from its sequence using physiochemical models. Recently, this has been exploited to computationally design DNA sequences encoding synthetic sRNA's - small RNA's which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.

This report will focus on one such system, introduced in \cite{Rodrigo2012}. It will extend existing understanding of the system beyond the qualitative, by proposing a quantitative model of gene expression, in the form of a set of ODE's, and fitting it to available time series data to estimate the model's unknown parameters.

>>>>

\section{Introduction}
\label{sec: Intro}
The process of gene expression can be summarised as follows: DNA is read, and a copy of it is made, in the form of an RNA molecule (this is called \textit{transcription}). This RNA molecule (known as messenger RNA, or mRNA) makes its way to a piece of cellular machinery called the Ribosome, which reads it, and makes a protein - which protein is made depends on the DNA sequence originally read (\textit{translation}).

 The path from genetic transcription to protein expression is naturally regulated in many ways \cite{MolecularBiology}. This regulation allows the cell to control protein expression, and so cell behaviour, in response to various environmental cues. The natural cell machinery which performs it takes the form a genetic circuits - networks of interacting gene expression regulators. This genetic circuitry offers rich possibilities for modification, and an important goal within synthetic biology is to understand and manipulate it.

As well as acting as the intermediate between DNA and protein, RNA molecules play direct and important roles in regulating gene expression \cite{Isaacs2006}. For the synthetic biologist looking to engineer regulation of genetic circuitry, RNA offers an attractive alternative to more traditional methods, which typically involve using proteins to regulate DNA transcription. In comparison to proteins, it is relatively straightforward to predict the structure and function of an RNA from its sequence using physiochemical models. Recently, this has been exploited to computationally design synthetic sRNA's - small RNA's which do not code for a protein, but rather have some direct regulatory function -  with regulatory behaviour that can be predicted \cite{Rodrigo2013}\cite{Rodrigo2012}.

This report will focus on one such sRNA system, introduced in \cite{Rodrigo2012}. It will extend existing understanding of it beyond the qualitative by proposing a quantitative model of gene expression, in the form of a set of ODE's, and fitting this model to available time series data to estimate its unknown parameters.
